---
title: "diff_gen"
author: "olly-mapps"
---

Contained in this notebook is the sandbox for code that I plan to use for a football scouting tool 


# Libraries
```{r}
library(tidyverse)
library(worldfootballR)
```

# Data Loading

We first define stat types to extract from worldsoccerR

```{r}
#Define stat types
stat_types = c("standard", "shooting", "passing", "passing_types", "gca", "defense", "possession", "playing_time", "misc", "keepers", "keepers_adv")
```

We next define a loop to extract these data based on the stat type

```{r}
#Define loop to iterate through the stat types
for (type in stat_types){
  
  #Load the data for the specific type
  temp <- data.frame(load_fb_big5_advanced_season_stats(stat_type = type, 
                                             team_or_player = "player"))
  #Print type to assess progress
  print(type)
  
  #Assign the data a suitable name
  assign(paste(type, "_stats", sep = ""), get("temp"))
}
```

# Data Wrangling

There are numeric columns that we don't need, so we drop them here. We also drop na values and convert the date to character for later

```{r}
#Define columns to leave out 
not_columns <- c("Age", "Born", "Prog", "A_minus_xA", "xAG", "A_minus_xAG", "PrgP")

#Create list of different types of stats scraped with worldsoccerR
type_list_raw <- list(standard_stats, shooting_stats, passing_stats, passing_types_stats, gca_stats, defense_stats, possession_stats, playing_time_stats, misc_stats, keepers_stats, keepers_adv_stats)

#Define a new list that cleans the data
type_list <- lapply(type_list_raw, function(x) x %>% 
                      
                      #Remove columns defined above
                      dplyr::select(-any_of(not_columns)) %>%
                      
                      #Removed na values
                      na.omit(.))

#Set the names of each element of the list to the stat type it represents
type_list <- setNames(type_list, stat_types)
```

# Converting to Stats

We next convert each column of each data-frame to percentiles, which we can imagine as standardized
'attributes'

```{r}
#Define for loop to cycle through each stat type in the list
for (type in stat_types){
  
  #Define for loop to cycle through each numeric value in the dataframe, apart from season year
  for (col in colnames(dplyr::select(type_list[[type]], where(is.numeric), -Season_End_Year))){
    
    #Calculate cumulative distribution percentile, and multiply by 100
    type_list[[type]][col] <- round(cume_dist(type_list[[type]][col])*100)
  }
}
```


# Finding Transfers 

We first filter to only include players who have had strictly more than 1 squad, this in effect filters only those players who have transferred.

```{r}
#Define for loop to cycle through each stat type
for (type in stat_types){
  
  #Overwrite each data frame with players with a number of distinct squads greater than 1
  type_list[[type]] <- type_list[[type]] %>% group_by(Player) %>% 
                        dplyr::filter(n_distinct(Squad) > 1) %>% 
                        ungroup()
}
```

# Calculating Stats

We next calculate the averages the stats over each players time at each club 

```{r}
#Define a for loop to cycle through each stat type
for (type in stat_types){
  
  #Overwrite each data frame with the means of each statistic during their time at each club
  type_list[[type]] <- type_list[[type]] %>% group_by(Player, Squad, Comp) %>% 
                        summarise(across(where(is.numeric), mean),
                          .groups = 'drop')
}
```

# Writing Stats

We first want to filter out the observations that lead to a transfer.

```{r}
for (type in stat_types){
  
  temp <- type_list[[type]] %>% 
          group_by(Player) %>%
          arrange(Season_End_Year) %>% 
          slice(1:(n()-1))
  
  assign(paste(type, "_from", sep = ""), get("temp"))
}
```

We write now write the data to csv

```{r}
write.csv(standard_from, "../Data/_from/standard_from.csv")
write.csv(shooting_from, "../Data/_from/shooting_from.csv")
write.csv(passing_from, "../Data/_from/passing_from.csv")
write.csv(passing_types_from, "../Data/_from/passing_types_from.csv")
write.csv(gca_from, "../Data/_from/gca_from.csv")
write.csv(defense_from, "../Data/_from/defense_from.csv")
write.csv(possession_from, "../Data/_from/possession_from.csv")
write.csv(playing_time_from, "../Data/_from/playing_time_from.csv")
write.csv(misc_from, "../Data/_from/misc_from.csv")
write.csv(keepers_from, "../Data/_from/keepers_from.csv")
write.csv(keepers_adv_from, "../Data/_from/keepers_adv_from.csv")
```

# Calculating Stats Differences

We next define a function that calculates the difference in percentiles between each transfer, and then produces a more manageable dataframe

```{r}
#Define function to generate new dataframe
transfer_gen <- function(x) {
  
  #Define empty dataframe to store stat changes between clubs
  squad_changes <- data.frame()
  
  #Sort by year so the differences are calculated in correct order
  x <- x %>% arrange(Season_End_Year)
  
  #For loop to cycle through each year
  for(i in 1:(nrow(x)-1)){
    
    #Define a temp variable to describe each pair of rows
    temp <- x[i:(i+1), ]
    
    #Flatten data frame to describe the change of squad and league
    temp <- temp %>% dplyr::select(!where(is.double)) %>% 
              rownames_to_column(var = "id") %>%
              pivot_wider(names_from = id, values_from = !id)
    
    #Merge the player columns since they will always be identical
    temp <- temp %>% mutate(Player = coalesce(Player_1, Player_2))
    temp <- temp %>% dplyr::select(-c("Player_1", "Player_2"))
    
    #Bind this players data to the overall changes dataframe
    squad_changes <- rbind(squad_changes, temp)
  }
  
  #Calculate differences between stats in each transfer
  stat_changes <- data.frame(diff(as.matrix(dplyr::select(x, where(is.numeric))), lag = 1))
  
  #Merge the two dataframes
  to_return <- data.frame(squad_changes, stat_changes)
  
  #Drop season year as it is no longer useful
  to_return <- to_return %>% dplyr::select(-Season_End_Year)
  
  #Return final data frame
  return(to_return)
}
```

We now apply this function to each data-set (Will take ages to run)

```{r}
#Define for loop to cycle through each stat type
for (type in stat_types){
  
  #Define empty data frame to hold stat changes for each stat type
  type_stat_changes <- data.frame()
  
  #Define list of unique players, so we can cycle through them in the for loop
  transferred_player_list <- as.list(unique(type_list[[type]]["Player"]))
  
  #Define for loop to cycle through each player
  for (player in transferred_player_list$Player){
    
    #Filter the stat dataframe to only the transfers concerning the given player
    player_info <- type_list[[type]] %>% dplyr::filter(type_list[[type]]["Player"] == player)
    
    #Bind the new dataframe to the stat changes data frame for this type
    type_stat_changes <- rbind(type_stat_changes, transfer_gen(player_info))
  
  }
  
  #Assign this new dataframe a suitable name
  assign(paste(type, "_changes", sep = ""), get("type_stat_changes"))
  
  #Print type to assess progress
  print(type)
  
}


```

## Write to csv

We finally write or dataframes to .csv files

```{r}
write.csv(standard_changes, "../Data/_changes/standard_changes.csv")
write.csv(shooting_changes, "../Data/_changes/shooting_changes.csv")
write.csv(passing_changes, "../Data/_changes/passing_changes.csv")
write.csv(passing_types_changes, "../Data/_from/passing_types_changes.csv")
write.csv(gca_changes, "../Data/_changes/gca_changes.csv")
write.csv(defense_changes, "../Data/_changes/defense_changes.csv")
write.csv(possession_changes, "../Data/_changes/possession_changes.csv")
write.csv(playing_time_changes, "../Data/_changes/playing_time_changes.csv")
write.csv(misc_changes, "../Data/_changes/misc_changes.csv")
write.csv(keepers_changes, "../Data/_changes/keepers_changes.csv")
write.csv(keepers_adv_changes, "../Data/_changes/keepers_adv_changes.csv")
```

